---
title: "SOA-PCR-Model"
author: "Nathan Ng"
date: '2022-04-10'
output: html_document
---

# PCR Model

Team: Part the bus 10/04/2022

## Training and Predicting using Tournament Dataset

#### Setting file directory and random seed

```{r}
setwd("C:/Users/natha/OneDrive/University/4th Year/ACTL4001/Group Project")
set.seed(8)
```

#### Loading Packages

```{r}
library(dplyr)
library(MASS)
library(readxl)
library(ggplot2)
library(ggfortify)
library(ISLR)
library(pls)
library(tidyr)
library(plyr)
```

#### Uploading of data and filtering of positions

```{r}
TournamentSet <- read_excel("Cleaned Dataset.xlsx", sheet = 12, n_max = 550)
FWSet <- TournamentSet[which(TournamentSet$Pos == "FW"),]
MWSet <- TournamentSet[which(TournamentSet$Pos == "MF"),]
DFSet <- TournamentSet[which(TournamentSet$Pos == "DF"),]
GKSet <- read_excel("Cleaned Dataset.xlsx", sheet = 13, n_max = 75)
```

#### Creating PCA plot to identify possible clusters

```{r}
Dataset <- read_excel("Cleaned Dataset.xlsx", sheet = 12, n_max = 550)
Dataset$Ranking <- floor(Dataset$`Nation Rank`/4)
Dataset_PCA <- prcomp(Dataset[,-c(1:3,26:28)], center = TRUE, scale. = TRUE)
summary(Dataset_PCA)
autoplot(Dataset_PCA, data = Dataset, colour = "Ranking", loadings = TRUE, loadings.label = TRUE, loadings.label.size = 3, 
         loadings.colour = "blue", loadings.label.colour = "blue")
```

#### Clean Non-numeric Variables for filtered datasets

```{r}
FWSetClean <- FWSet[,-c(1:3,5,26,27)]
MWSetClean <- MWSet[,-c(1:3,5,26,27)]
DFSetClean <- DFSet[,-c(1:3,5,26,27)]
TournamentSetClean <- TournamentSet[,-c(1:3,5,26,27)] 
```

### Modelling Entire Tournament Dataset

### Split data into a training and test set

```{r}
sample <- sample(nrow(TournamentSet),nrow(TournamentSet)*0.7)
trainset <- TournamentSet[sample,]
testset <- TournamentSet[-sample,]
```

#### Removing non-numerical data

```{r}
train <- trainset[,-c(1:3,5,26,27)]
test <- testset[,-c(1:3,5,26,27)]
```

#### Applying PCR to create model using training dataset

```{r}
pcrmodel <- pcr(train$`Nation Rank` ~. , data = train, scale = TRUE, validation = "CV")
```

#### Validation plot to identify best number of principle components

```{r}
validationplot(pcrmodel, val.type = "MSE")
```

#### Creating predictor vectors for training and testing set

```{r}
x_train <- model.matrix(train$`Nation Rank` ~., train)[,-1]
x_test <- model.matrix(test$`Nation Rank` ~., test)[,-1]
```

#### Creating response vectors from training and testing set

```{r}
y_train <- as.numeric(unlist(train[,22]))
y_test <- as.numeric(unlist(test[,22]))
```

#### Predicting training and testing sets with PCR model

```{r}
pcrtrainpredict <- predict(pcrmodel, x_train, ncomp = 7)
pcrpredict <- predict(pcrmodel, x_test, ncomp = 7)
```

#### Computing the MSE of model

```{r}
mean((pcrpredict-y_test)^2)
```

#### Tabulating results for both train and test datasets

```{r}
TournamentTrain <- data.frame(trainset$Nation,y_train,pcrtrainpredict)
colnames(TournamentTrain) <- c("Nation", "True Value", "Predicted Value")
aggregate(TournamentTrain[,2:3], list(TournamentTrain$Nation), mean)


TournamentTest <- data.frame(testset$Nation,y_test,pcrpredict)
colnames(TournamentTest) <- c("Nation", "True Value", "Predicted Value")
aggregate(TournamentTest[,2:3], list(TournamentTest$Nation), mean)
```

### PCR Model for Forward Players

#### Split data into a training and test set

```{r}
sampleF <- sample(nrow(FWSet),nrow(FWSet)*0.7)
trainsetF <- FWSet[sampleF,]
testsetF <- FWSet[-sampleF,]
```

#### Removing non-numerical data

```{r}
trainF <- trainsetF[,-c(1:3,5,26,27)]
testF <- testsetF[,-c(1:3,5,26,27)]
```

#### Applying PCR to create model using training dataset

```{r}
pcrmodelF <- pcr(trainF$`Nation Rank` ~. , data = trainF, scale = TRUE, validation = "CV")
```

#### Validation plot to identify best number of principle components

```{r}
validationplot(pcrmodelF, val.type = "MSE")
```

#### Creating predictor vectors for training and testing set

```{r}
x_trainF <- model.matrix(trainF$`Nation Rank` ~., trainF)[,-1]
x_testF <- model.matrix(testF$`Nation Rank` ~., testF)[,-1]
```

#### Creating response vectors from training and testing set

```{r}
y_trainF <- as.numeric(unlist(trainF[,22]))
y_testF <- as.numeric(unlist(testF[,22]))
```

#### Predicting training and testing sets with PCR model

```{r}
pcrtrainpredictF <- predict(pcrmodelF, x_trainF, ncomp = 6)
pcrpredictF <- predict(pcrmodelF, x_testF, ncomp = 6)
```

#### Computing the MSE of model

```{r}
mean((pcrpredictF-y_testF)^2)
```

#### Tabulating results for both train and test datasets and identifying MSE

```{r}
TournamentTrainF <- data.frame(trainsetF$Nation,y_trainF,pcrtrainpredictF)
colnames(TournamentTrainF) <- c("Nation", "True Value", "Predicted Value")
TourTrainF <- aggregate(TournamentTrainF[,2:3], list(TournamentTrainF$Nation), mean)
mean((TourTrainF$`True Value`-TourTrainF$`Predicted Value`)^2)

TournamentTestF <- data.frame(testsetF$Nation,y_testF,pcrpredictF)
colnames(TournamentTestF) <- c("Nation", "True Value", "Predicted Value")
TourTestF <-aggregate(TournamentTestF[,2:3], list(TournamentTestF$Nation), mean)
mean((TourTestF$`True Value`-TourTestF$`Predicted Value`)^2)
```

### PCR Model for Midfield Players

#### Split data into a training and test set

```{r}
sampleM <- sample(nrow(MWSet),nrow(MWSet)*0.7)
trainsetM <- MWSet[sampleM,]
testsetM <- MWSet[-sampleM,]
```

#### Removing non-numerical data

```{r}
trainM <- trainsetM[,-c(1:3,5,26,27)]
testM <- testsetM[,-c(1:3,5,26,27)]
```

#### Applying PCR to create model using training dataset

```{r}
pcrmodelM <- pcr(trainM$`Nation Rank` ~. , data = trainM, scale = TRUE, validation = "CV")
```

#### Validation plot to identify best number of principle components

```{r}
validationplot(pcrmodelM, val.type = "MSE")
```

#### Creating predictor vectors for training and testing set

```{r}
x_trainM <- model.matrix(trainM$`Nation Rank` ~., trainM)[,-1]
x_testM <- model.matrix(testM$`Nation Rank` ~., testM)[,-1]
```

#### Creating response vectors from training and testing set

```{r}
y_trainM <- as.numeric(unlist(trainM[,22]))
y_testM <- as.numeric(unlist(testM[,22]))
```

#### Predicting training and testing sets with PCR model

```{r}
pcrtrainpredictM <- predict(pcrmodelM, x_trainM, ncomp = 5)
pcrpredictM <- predict(pcrmodelM, x_testM, ncomp = 5)
```

#### Computing the MSE of model

```{r}
mean((pcrpredictM-y_testM)^2)
```

#### Tabulating results for both train and test datasets and identifying MSE

```{r}
TournamentTrainM <- data.frame(trainsetM$Nation,y_trainM,pcrtrainpredictM)
colnames(TournamentTrainM) <- c("Nation", "True Value", "Predicted Value")
TourTrainM <- aggregate(TournamentTrainM[,2:3], list(TournamentTrainM$Nation), mean)
mean((TourTrainM$`True Value`-TourTrainM$`Predicted Value`)^2)

TournamentTestM <- data.frame(testsetM$Nation,y_testM,pcrpredictM)
colnames(TournamentTestM) <- c("Nation", "True Value", "Predicted Value")
TourTestM <-aggregate(TournamentTestM[,2:3], list(TournamentTestM$Nation), mean)
mean((TourTestM$`True Value`-TourTestM$`Predicted Value`)^2)
```

### PCR Model for Defense Players

#### Split data into a training and test set

```{r}
sampleD <- sample(nrow(DFSet),nrow(DFSet)*0.7)
trainsetD <- DFSet[sampleD,]
testsetD <- DFSet[-sampleD,]
```

#### Removing non-numerical data

```{r}
trainD <- trainsetD[,-c(1:3,5,26,27)]
testD <- testsetD[,-c(1:3,5,26,27)]
```

#### Applying PCR to create model using training dataset

```{r}
pcrmodelD <- pcr(trainD$`Nation Rank` ~. , data = trainD, scale = TRUE, validation = "CV")
```

#### Validation plot to identify best number of principle components

```{r}
validationplot(pcrmodelD, val.type = "MSE")
```

#### Creating predictor vectors for training and testing set

```{r}
x_trainD <- model.matrix(trainD$`Nation Rank` ~., trainD)[,-1]
x_testD <- model.matrix(testD$`Nation Rank` ~., testD)[,-1]
```

#### Creating response vectors from training and testing set

```{r}
y_trainD <- as.numeric(unlist(trainD[,22]))
y_testD <- as.numeric(unlist(testD[,22]))
```

#### Predicting training and testing sets with PCR model

```{r}
pcrtrainpredictD <- predict(pcrmodelD, x_trainD, ncomp = 3)
pcrpredictD <- predict(pcrmodelD, x_testD, ncomp = 3)
```

#### Computing the MSE of model

```{r}
mean((pcrpredictD-y_testD)^2)
```

#### Tabulating results for both train and test datasets and identifying MSE

```{r}
TournamentTrainD <- data.frame(trainsetD$Nation,y_trainD,pcrtrainpredictD)
colnames(TournamentTrainD) <- c("Nation", "True Value", "Predicted Value")
TourTrainD <- aggregate(TournamentTrainD[,2:3], list(TournamentTrainD$Nation), mean)
mean((TourTrainD$`True Value`-TourTrainD$`Predicted Value`)^2)

TournamentTestD <- data.frame(testsetD$Nation,y_testD,pcrpredictD)
colnames(TournamentTestD) <- c("Nation", "True Value", "Predicted Value")
TourTestD <-aggregate(TournamentTestD[,2:3], list(TournamentTestD$Nation), mean)
mean((TourTestD$`True Value`-TourTestD$`Predicted Value`)^2)
```

## PCR Model for Goalkeeping Players

#### Split data into a training and test set

```{r}
sampleG <- sample(nrow(GKSet),nrow(GKSet)*0.7)
trainsetG <- GKSet[sampleG,]
testsetG <- GKSet[-sampleG,]
```

#### Removing non-numerical data

```{r}
trainG <- trainsetG[,-c(1:5)]
testG <- testsetG[,-c(1:5)]
```

#### Applying PCR to create model using training dataset

```{r}
pcrmodelG <- pcr(trainG$`Nation Rank` ~. , data = trainG, scale = TRUE, validation = "CV")
```

#### Validation plot to identify best number of principle components

```{r}
validationplot(pcrmodelG, val.type = "MSE")
```

#### Creating predictor vectors for training and testing set

```{r}
x_trainG <- model.matrix(trainG$`Nation Rank` ~., trainG)[,-1]
x_testG <- model.matrix(testG$`Nation Rank` ~., testG)[,-1]
```

#### Creating response vectors from training and testing set

```{r}
y_trainG <- as.numeric(unlist(trainG[,7]))
y_testG <- as.numeric(unlist(testG[,7]))
```

#### Predicting training and testing sets with PCR model

```{r}
pcrtrainpredictG <- predict(pcrmodelG, x_trainG, ncomp = 5)
pcrpredictG <- predict(pcrmodelG, x_testG, ncomp = 5)
```

#### Computing the MSE of model

```{r}
mean((pcrpredictG-y_testG)^2)
```

#### Tabulating results for both train and test datasets and identifying MSE

```{r}
TournamentTrainG <- data.frame(trainsetG$Nation,y_trainG,pcrtrainpredictG)
colnames(TournamentTrainG) <- c("Nation", "True Value", "Predicted Value")
TourTrainG <- aggregate(TournamentTrainG[,2:3], list(TournamentTrainG$Nation), mean)
mean((TourTrainG$`True Value`-TourTrainG$`Predicted Value`)^2)

TournamentTestG <- data.frame(testsetG$Nation, y_testG,pcrpredictG)
colnames(TournamentTestG) <- c("Nation", "True Value", "Predicted Value")
TourTestG <-aggregate(TournamentTestG[,2:3], list(TournamentTestG$Nation), mean)
mean((TourTestG$`True Value`-TourTestG$`Predicted Value`)^2)
```

## Results

#### Creating results table for training dataset

```{r}
df <- merge(TourTrainD, TourTrainF, by = "Group.1", all = TRUE)
colnames(df) <- c("Group.1",2,3,4)
df <- merge(df, TourTrainG, by = "Group.1", all = TRUE)
df <- merge(df, TourTrainM, by = "Group.1", all = TRUE)
df<- df[-c(4,6,8)]
colnames(df) <- c("Nation", "True Rank", "D", "F", "G", "M")
df$AggregateMean <- rowMeans(df[,c(3:6)], na.rm = TRUE)
mean((df$`True Rank`-df$AggregateMean)^2)
```

#### Creating results table for test dataset

```{r}
df2 <- merge(TourTestD, TourTestF, by = "Group.1", all = TRUE)
colnames(df2) <- c("Group.1",2,3,4)
df2 <- merge(df2, TourTestG, by = "Group.1", all = TRUE)
df2 <- merge(df2, TourTestM, by = "Group.1", all = TRUE)
df2<- df2[-c(4,6,8)]
colnames(df2) <- c("Nation", "True Rank", "D", "F", "G", "M")
df2$AggregateMean <- rowMeans(df2[,c(3:6)], na.rm = TRUE)
df2[20,2]<- 5
mean((df2$`True Rank`-df2$AggregateMean)^2)
```
